---
title: "VPKSR_Survey_PhilHerman"
author: "RobertChapman"
date: "2023-12-12"
output:
  html_document:
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(data.table)
library(DT)

data=list("SRCombined"=read.csv("SR_12-11.csv"),"VPKCombined"=read.csv("VPK_12-11.csv"))


PlotTabulate=function(var, data, subpath=""){
  sapply(names(data), simplify=FALSE, USE.NAMES = TRUE, function(d){
    # print(paste0(subpath,d,"_",var,".png"))
    # print(var %in% names(data[[d]]))
     if(var %in% names(data[[d]])){
      t=table(data[[d]][,var], useNA="always")
      tp=prop.table(t)
      write.csv(rbind("Freq"=t,"Prop"=tp),paste0(subpath,d,"_",var,"_table.csv"))
      png(paste0(subpath,d,"_",var,".png"))
      names(t)[is.na(names(t))]="Missing/NA"
      for(name in which(nchar(names(t))>20)){
        loc=unlist(gregexpr("\\s",names(t)[name]))[unlist(gregexpr("\\s",names(t)[name]))>12][1]
        substr(names(t)[name], loc, loc) <- "\n"}
      
      #par(mar = c(bottom, left, top, right))
        par(mar = c(12.1, 4.1, 4.1, 2.1))
        par(las=2)
        b=barplot(t, ylim=c(0,max(t)*1.2))
        text(b, t + max(t)*0.1, t)
      dev.off()

      out=list("plot"=b,"freqtable"=t,"proptable"=tp)
       }
     else{out=list(NA)}
  })
}



```


### Analyis plan from Phil 12/4
#### SR

  *Analysis 1: Price of Care a prohibitive factor* <- CTree <p>
      Outcome variable: Childcare too expensive, coded as dichotomous (either too expensive or not) <br>
      Predictors:  Marital status, (combining married & living with partners), income, # of children needing care, preferred amount of money, transportation, ages of children needing care (collapsing across all ages - one value per row), parent education
  <p>
  *Analysis 2: Cross tabs/descriptives* <br>
      1) Preferred distance by mode of transportation and people who did/didnt endorse pref for ‘conveniently located’<br>
```{r SRA2Xtab1}
    DistancebyTransModeforNotConvLocatedTrue=table(
        data$SRCombined$PrimaryTransportationMode[
          grep("*conveniently located*",data$SRCombined$TroubleFindingChildcare, invert=FALSE)], 
        data$SRCombined$PreferredTravelDistance[
          grep("*conveniently located*",data$SRCombined$TroubleFindingChildcare, invert=FALSE)])
      write.csv(DistancebyTransModeforNotConvLocatedTrue, "SR_xtabs_PrefDistbyTranspMode_NotConvLocTRUE.csv")  

    #Distance by Transportation Mode table for Not Conveniently Located = True
      datatable(matrix(DistancebyTransModeforNotConvLocatedTrue, ncol = ncol(DistancebyTransModeforNotConvLocatedTrue), dimnames = dimnames(DistancebyTransModeforNotConvLocatedTrue)), caption="Distance by Transportation Mode table for Not Conveniently Located = True")
      
      DistancebyTransModeforNotConvLocatedFalse=table(
        data$SRCombined$PrimaryTransportationMode[
          !grepl("*conveniently located*",data$SRCombined$TroubleFindingChildcare)], 
        data$SRCombined$PreferredTravelDistance[
          !grepl("*conveniently located*",data$SRCombined$TroubleFindingChildcare)])
      
         write.csv(DistancebyTransModeforNotConvLocatedFalse, "SR_xtabs_PrefDistbyTranspMode_NotConvLocFALSE.csv")
         
      #Distance by Transportation Mode table for Not Conveniently Located = False   
      datatable(matrix(DistancebyTransModeforNotConvLocatedFalse, ncol = ncol(DistancebyTransModeforNotConvLocatedFalse), dimnames = dimnames(DistancebyTransModeforNotConvLocatedFalse)), caption="Distance by Transportation Mode table for Not Conveniently Located = False")
        
```
        
    2) Provider quality with parent education<br>
```{r SRA2Xtab2}

NotQualitybyEducation=table(data$SRCombined$Education, "Provider Not High Quality"=grepl("*quality*",data$SRCombined$TroubleFindingChildcare))

write.csv(NotQualitybyEducation,"SR_xtabs_EducationbyProviderNotQuality.csv")

#Education by Provider Not High Quality   
      datatable(matrix(NotQualitybyEducation, ncol = ncol(NotQualitybyEducation), dimnames = dimnames(NotQualitybyEducation)), caption="Education by Provider Not High Quality")


```
    3) "Provider not open during hours I need" vs Ranges for open/closed preferences<br>
```{r SRA2Xtab3}


TotalNotOpenHoursNeededbyBusinessHours=table("Provider Not Open Hours I need"=grepl("*hours*",data$SRCombined$TroubleFindingChildcare),
"Preferred Hours Outside 7a-6p"=(data$SRCombined$ProviderOpensAM<"7:00"|data$SRCombined$ProviderClosesPM>"6:00"))

TotalNotOpenHoursNeededbyBusinessHours

ProviderHoursbyNotOpenHoursNeededTrue=table("Opens AM"=data$SRCombined$ProviderOpensAM[grep("*hours*",data$SRCombined$TroubleFindingChildcare)],
      "Closes PM"=data$SRCombined$ProviderClosesPM[grep("*hours*",data$SRCombined$TroubleFindingChildcare)])

write.csv(ProviderHoursbyNotOpenHoursNeededTrue,"SRxtabs_PrefHoursbyProvNotOpenHours.csv")

#Provider Hours by Not Open Hours Needed = True   
      datatable(matrix(ProviderHoursbyNotOpenHoursNeededTrue, ncol = ncol(ProviderHoursbyNotOpenHoursNeededTrue), dimnames = dimnames(ProviderHoursbyNotOpenHoursNeededTrue)), caption="Provider Hours by Not Open Hours Needed = True ", options = list(pageLength = 25))

ProviderHoursbyNotOpenHoursNeededFalse=table("Opens AM"=data$SRCombined$ProviderOpensAM[grep("*hours*",data$SRCombined$TroubleFindingChildcare, invert=TRUE)],
      "Closes PM"=data$SRCombined$ProviderClosesPM[grep("*hours*",data$SRCombined$TroubleFindingChildcare, invert=TRUE)])

write.csv(ProviderHoursbyNotOpenHoursNeededFalse,"SRxtabs_PrefHoursbyProvOpenHours.csv")

#Provider Hours by Not Open Hours Needed = False   
      datatable(matrix(ProviderHoursbyNotOpenHoursNeededFalse, ncol = ncol(ProviderHoursbyNotOpenHoursNeededFalse), dimnames = dimnames(ProviderHoursbyNotOpenHoursNeededFalse)), caption="Provider Hours by Not Open Hours Needed = False ", options = list(pageLength = 25))


```
    4) Ranges for open/closed preferences <- general for everyone <br>
```{r SRA2Xtab4}

GenProviderHours=table("Opens AM"=data$SRCombined$ProviderOpensAM,
      "Closes PM"=data$SRCombined$ProviderClosesPM)

write.csv(GenProviderHours,"SRxtabs_PrefHours.csv")

#General Provider Hours   
      datatable(matrix(GenProviderHours, ncol = ncol(GenProviderHours), dimnames = dimnames(GenProviderHours)), caption="General Provider Hours", options = list(pageLength = 25))

```

#### VPK:
    <p>
    *No longer needed care*<br>
    **Section 1: Sent kids to stay with family/friends**<p>
        Household Size distribution<br>
        Income distribution<br>
        marital status frequencies<br>
        what responses do they select for reasons having trouble finding a provider<p>
```{r VPKA1, results='hide'}

Sect1Indx=grep("*I decided to have my child stay with family/friends*",data$VPKCombined$WhyNoLongerNeeded)

    VPKA1_HH_Size=PlotTabulate("HouseHoldSize", list("VPK_Child Staying with FamilyFriends"=data$VPKCombined[Sect1Indx,]))
    barplot(VPKA1_HH_Size[[1]]$freqtable,main="VPK_Child Staying with FamilyFriends \n by House Hold Size")
    VPKA1_HH_Inc=PlotTabulate("HouseHoldIncome", list("VPK_Child Staying with FamilyFriends"=data$VPKCombined[Sect1Indx,]))
    par(mar = c(12.1, 4.1, 4.1, 2.1))
    par(las=2)
    barplot(VPKA1_HH_Inc[[1]]$freqtable, main="VPK_Child Staying with FamilyFriends \n by House Hold Income")
    VPKA1_Marital=PlotTabulate("MaritalStatus", list("VPK_Child Staying with FamilyFriends"=data$VPKCombined[Sect1Indx,]))
    par(mar = c(12.1, 4.1, 4.1, 2.1))
    par(las=2)
    barplot(VPKA1_Marital[[1]]$freqtable, main="VPK_Child Staying with FamilyFriends \n by Marital Status")
    
    png("FamFriends_Trouble.png")
    par(mar = c(10.1, 4.1, 4.1, 2.1))
    par(las=2)
    barplot(apply(data$VPKCombined[Sect1Indx,grep("^Trbl",names(data$VPKCombined))], 2, function(c){sum(as.numeric(c),na.rm=TRUE)})) 
    dev.off()

    par(mar = c(10.1, 4.1, 4.1, 2.1))
    par(las=2)
    barplot(apply(data$VPKCombined[Sect1Indx,grep("^Trbl",names(data$VPKCombined))], 2, function(c){sum(as.numeric(c),na.rm=TRUE)}), main="VPK_Child Staying with FamilyFriends\nTrouble Finding a Provider") 
    
    
```
    
  **Section 2: Stay at home with child**<p>
      Household Size distribution<br>
      Income distribution<br>
      marital status frequencies<br>
      what responses do they select for reasons having trouble finding a provider<br>
```{r VPKA2, results='hide'}

Sect2Indx=grep("*I decided to stay home with my child*",data$VPKCombined$WhyNoLongerNeeded)

    VPKA2_HH_Size=PlotTabulate("HouseHoldSize", list("VPK_Child Staying Home"=data$VPKCombined[Sect2Indx,]))
    barplot(VPKA2_HH_Size[[1]]$freqtable,main="VPK_Child Staying Home \n by House Hold Size")
    VPKA2_HH_Inc=PlotTabulate("HouseHoldIncome", list("VPK_Child Staying Home"=data$VPKCombined[Sect2Indx,]))
    par(mar = c(12.1, 4.1, 4.1, 2.1))
    par(las=2)
    barplot(VPKA2_HH_Inc[[1]]$freqtable, main="VPK_Child Staying Home \n by House Hold Income")
    VPKA2_Marital=PlotTabulate("MaritalStatus", list("VPK_Child Staying Home"=data$VPKCombined[Sect2Indx,]))
    par(mar = c(12.1, 4.1, 4.1, 2.1))
    par(las=2)
    barplot(VPKA2_Marital[[1]]$freqtable, main="VPK_Child Staying Home \n by Marital Status")
    
    png("FamFriends_Trouble.png")
    par(mar = c(10.1, 4.1, 4.1, 2.1))
    par(las=2)
    barplot(apply(data$VPKCombined[Sect2Indx,grep("^Trbl",names(data$VPKCombined))], 2, function(c){sum(as.numeric(c),na.rm=TRUE)})) 
    dev.off()

    par(mar = c(10.1, 4.1, 4.1, 2.1))
    par(las=2)
    barplot(apply(data$VPKCombined[Sect2Indx,grep("^Trbl",names(data$VPKCombined))], 2, function(c){sum(as.numeric(c),na.rm=TRUE)}), main="VPK_Child Staying Home\nTrouble Finding a Provider") 
    
    
```    

  **Section 3: Alternative care**<p>
      Household Size distribution<br>
      Income distribution<br>
      marital status frequencies<br>
      what responses do they select for reasons having trouble finding a provider<br>

```{r VPKA3, results='hide'}

Sect3Indx=grep("*alternative*",data$VPKCombined$WhyNoLongerNeeded)

    VPKA3_HH_Size=PlotTabulate("HouseHoldSize", list("VPK_AlternativeCare"=data$VPKCombined[Sect3Indx,]))
    barplot(VPKA3_HH_Size[[1]]$freqtable,main="VPK_AlternativeCare \n by House Hold Size")
    VPKA3_HH_Inc=PlotTabulate("HouseHoldIncome", list("VPK_AlternativeCare"=data$VPKCombined[Sect3Indx,]))
    par(mar = c(12.1, 4.1, 4.1, 2.1))
    par(las=2)
    barplot(VPKA3_HH_Inc[[1]]$freqtable, main="VPK_AlternativeCare \n by House Hold Income")
    VPKA3_Marital=PlotTabulate("MaritalStatus", list("VPK_AlternativeCare"=data$VPKCombined[Sect3Indx,]))
    par(mar = c(12.1, 4.1, 4.1, 2.1))
    par(las=2)
    barplot(VPKA3_Marital[[1]]$freqtable, main="VPK_AlternativeCare \n by Marital Status")
    
    png("FamFriends_Trouble.png")
    par(mar = c(10.1, 4.1, 4.1, 2.1))
    par(las=2)
    barplot(apply(data$VPKCombined[Sect3Indx,grep("^Trbl",names(data$VPKCombined))], 2, function(c){sum(as.numeric(c),na.rm=TRUE)})) 
    dev.off()

    par(mar = c(10.1, 4.1, 4.1, 2.1))
    par(las=2)
    barplot(apply(data$VPKCombined[Sect3Indx,grep("^Trbl",names(data$VPKCombined))], 2, function(c){sum(as.numeric(c),na.rm=TRUE)}), main="VPK_AlternativeCare\nTrouble Finding a Provider") 
    
    
``` 

### Discussion with Herman and Phil, 12/5
#### Additional analyses

      Adjusted income:
      Income/# of folks in household-> (range)
      
```{r AdjustedIncome}

data$SRCombined$AdjHouseHoldIncome=apply(data$SRCombined, 1, function(row){

  if(!any(is.na(row[c("HouseHoldIncome","HouseHoldSize")]))){
    
    inc=switch(row["HouseHoldIncome"],
      "Less than $14,000"=14000,
      "$100,000 or more"=100000,
      "$15,000 to 19,000"=15000,
      "$20,000 to 29,000"=20000,
      "$30,000 to $39,000"=30000,
      "$40,000 to $49,000"=40000,
      "$50,000 to $74,000"=50000,
      "$75,000 to $100,000"=75000,
      "Prefer not to answer"=NA)

    size=switch(row["HouseHoldSize"],
      "1"=1,
      "2"=2,
      "3"=3,
      "4"=4,
      "5"=5,
      "6+"=6)
  
    out=as.numeric(inc/size)
  }else{out=NA}
  out
})

hist(data$SRCombined$AdjHouseHoldIncome)
summary(data$SRCombined$AdjHouseHoldIncome)

data$VPKCombined$AdjHouseHoldIncome=apply(data$VPKCombined, 1, function(row){

  if(!any(is.na(row[c("HouseHoldIncome","HouseHoldSize")]))){
    
    inc=switch(row["HouseHoldIncome"],
      "Less than $14,000"=14000,
      "$100,000 or more"=100000,
      "$15,000 to 19,000"=15000,
      "$20,000 to 29,000"=20000,
      "$30,000 to $39,000"=30000,
      "$40,000 to $49,000"=40000,
      "$50,000 to $74,000"=50000,
      "$75,000 to $100,000"=75000,
      "Prefer not to answer"=NA)

    size=switch(row["HouseHoldSize"],
      "1"=1,
      "2"=2,
      "3"=3,
      "4"=4,
      "5"=5,
      "6+"=6)
  
    out=as.numeric(inc/size)
  }else{out=NA}
  out
})

hist(data$VPKCombined$AdjHouseHoldIncome)
summary(data$VPKCombined$AdjHouseHoldIncome)

```
  
      
      Calc -> adjusted to money willing to spend per child (adj range)
      Price per kid

```{r AdjustedWillingToSpend}

data$SRCombined$AdjWillingToSpend=apply(data$SRCombined, 1, function(row){

  if(!any(is.na(row[c("WillingToSpend","HouseHoldSize")]))){

    spend=switch(unlist(row["WillingToSpend"]),
      "$0-25"=0,
      "$25-50"=25,
      "$50-75"=50,
      "$75-100"=75,
      "$100-150"=100,
      "Greater than $150"=150)

    size=switch(unlist(row["HouseHoldSize"]),
      "1"=1,
      "2"=2,
      "3"=3,
      "4"=4,
      "5"=5,
      "6+"=6)
  
    out=as.numeric(spend/size)
  }else{out=NA}
  out
})

hist(data$SRCombined$AdjWillingToSpend)
summary(data$SRCombined$AdjWillingToSpend)


data$VPKCombined$AdjWillingToSpend=apply(data$VPKCombined, 1, function(row){

  if(!any(is.na(row[c("WillingToSpend","HouseHoldSize")]))){

    spend=switch(unlist(row["WillingToSpend"]),
      "$0-25"=0,
      "$25-50"=25,
      "$50-75"=50,
      "$75-100"=75,
      "$100-150"=100,
      "Greater than $150"=150)

    size=switch(unlist(row["HouseHoldSize"]),
      "1"=1,
      "2"=2,
      "3"=3,
      "4"=4,
      "5"=5,
      "6+"=6)
  
    out=as.numeric(spend/size)
  }else{out=NA}
  out
})

hist(data$VPKCombined$AdjWillingToSpend)
summary(data$VPKCombined$AdjWillingToSpend)


```      
      
      
      
      
      Dig deeper on categories for specific reasons for trouble finding providers (try to identify characteristics)
      
      Trees -> VPK Outcome (Other, Didn’t Experience Challenges, Expensive, public)
      
```{r ctree, class.source = 'fold-show'}

      # Trees -> VPK 
      #   Outcome (Other, Didn’t Experience Challenges, Expensive, Public)
            table(data$VPKCombined$Trbl_NoChlgns)
            table(data$VPKCombined$Trbl_Expnsv)
            table(data$VPKCombined$Trbl_NoPubSchls)
            table(data$VPKCombined$Trbl_Other)

      #   Predictors:  
            #Marital status
              table(data$VPKCombined$MaritalStatus)
            #income & adjincome, 
              table(data$VPKCombined$HouseHoldIncome)
              table(data$VPKCombined$AdjHouseHoldIncome)
            # of children needing care-> we have # of kids entering KG and HH Size...
              table(data$VPKCombined$NumChildrenEnteringKG)
            #preferred amount of money, adjusted
              table(data$VPKCombined$WillingToSpend)
              table(data$VPKCombined$AdjWillingToSpend)
            #transportation, 
              table(data$VPKCombined$PrimaryTransportationMode)
            #ages of children needing care (collapsing across all ages - one value per row),
              #not in VPK, only in SR
            #parent education
              table(data$VPKCombined$Education)
          
              
data$VPKCombined$PrimaryTransportationMode[data$VPKCombined$PrimaryTransportationMode=="Walking"]="Walking/Bicycle"
data$VPKCombined$PrimaryTransportationMode[data$VPKCombined$PrimaryTransportationMode=="Bicycle"]="Walking/Bicycle"

data$VPKCombined$MaritalStatus[data$VPKCombined$MaritalStatus=="Living with partner in same household"]="Living with partner in same household/Married"
data$VPKCombined$MaritalStatus[data$VPKCombined$MaritalStatus=="Married"]="Living with partner in same household/Married"

data$VPKCombined$MaritalStatus[data$VPKCombined$MaritalStatus=="Single"]="Widowed/Divorced/Separated/Single"
data$VPKCombined$MaritalStatus[data$VPKCombined$MaritalStatus=="Divorced/Separated"]="Widowed/Divorced/Separated/Single"
data$VPKCombined$MaritalStatus[data$VPKCombined$MaritalStatus=="Widowed"]="Widowed/Divorced/Separated/Single"

data$VPKCombined$NumAdultsHH[data$VPKCombined$MaritalStatus=="Living with partner in same household/Married"]=2
data$VPKCombined$NumAdultsHH[data$VPKCombined$MaritalStatus=="Widowed/Divorced/Separated/Single"]=1



data$VPKCombined$AdultAdjHouseHoldIncome=apply(data$VPKCombined, 1, function(row){

  if(!any(is.na(row[c("HouseHoldIncome","NumAdultsHH")]))){
    
    inc=switch(row["HouseHoldIncome"],
      "Less than $14,000"=14000,
      "$100,000 or more"=100000,
      "$15,000 to 19,000"=15000,
      "$20,000 to 29,000"=20000,
      "$30,000 to $39,000"=30000,
      "$40,000 to $49,000"=40000,
      "$50,000 to $74,000"=50000,
      "$75,000 to $100,000"=75000,
      "Prefer not to answer"=NA)

  
    out=as.numeric(as.numeric(inc)/as.numeric(row["NumAdultsHH"]))
  }else{out=NA}
  out
})
              
              
              
              
              
              
d=data$VPKCombined
  for(c in names(d)){
      if(class(d[,c])=="character"){d[,c]=as.factor(d[,c])}
      if(grepl("Trbl",c)){d[,c]=as.factor(d[,c])}   
      }

for(v in c("Trbl_NoChlgns", "Trbl_Expnsv", "Trbl_NoPubSchls", "Trbl_Other", "MaritalStatus", "HouseHoldIncome", "AdjHouseHoldIncome", "NumChildrenEnteringKG", "WillingToSpend", "AdjWillingToSpend", "PrimaryTransportationMode", "Education")){
  print(paste0("Variable: ",v,", format -> ",class(d[,v])))
  
}




set.seed(27)
#install.packages("partykit")
library(partykit)
#install.packages("party")
library(party)


# 
# mod.VPK=ctree(Trbl_NoChlgns + Trbl_Expnsv + Trbl_NoPubSchls + Trbl_Other ~  MaritalStatus + HouseHoldIncome + NumChildrenEnteringKG + WillingToSpend2 + PrimaryTransportationMode + Education,
#                    data=d,
#                controls=ctree_control(minbucket=.01*nrow(d)))
# png("out.png",1200,1200)
# plot(mod.VPK)
# dev.off()
# 

d$WillingToSpend2=unlist(sapply(d$WillingToSpend, function(x){
  if(!is.na(x)){
    out=switch(
          as.character(x),
          "$0-25"=0 ,
            "$25-50"=1,
          "$50-75"=2,
          "$75-100"=3,
          "$100-150"=4,
          "Greater than $150"=5)}
    else{out=NA}
    out
    }))
# 
# ####
# mod.VPK2=ctree(Trbl_NoPubSchls ~  MaritalStatus + HouseHoldIncome + NumChildrenEnteringKG + WillingToSpend2 + PrimaryTransportationMode + Education,
#                    data=d,
#                controls=ctree_control(minbucket=.01*nrow(d)))
# 
# ###
# ####
# mod.VPK3=ctree(Trbl_NoPubSchls ~  MaritalStatus + HouseHoldIncome + NumChildrenEnteringKG + WillingToSpend2 + PrimaryTransportationMode,
#                    data=d,
#                controls=ctree_control(minbucket=.01*nrow(d)))
# 
# ####
# mod.VPK4=ctree(Trbl_NoPubSchls ~  MaritalStatus + AdultAdjHouseHoldIncome + NumChildrenEnteringKG + WillingToSpend2 + PrimaryTransportationMode,
#                    data=d,
#                controls=ctree_control(minbucket=.01*nrow(d)))
# 
# ####
# mod.VPK5=ctree(Trbl_NoPubSchls ~  MaritalStatus + HouseHoldIncome + NumChildrenEnteringKG + WillingToSpend2 + PrimaryTransportationMode,
#                    data=d,
#                controls=ctree_control(minbucket=.01*nrow(d)))
# 
# plot(mod.VPK5)
# 
# ####
# mod.VPK6=ctree(Trbl_Expnsv ~  MaritalStatus + HouseHoldIncome + NumChildrenEnteringKG + WillingToSpend2 + PrimaryTransportationMode,
#                    data=d,
#                controls=ctree_control(minbucket=.01*nrow(d)))
# 
# plot(mod.VPK6)
# 
# ####
# mod.VPK7=ctree(Trbl_NoChlgns ~  MaritalStatus + HouseHoldIncome + NumChildrenEnteringKG + WillingToSpend2 + PrimaryTransportationMode + Why_AlternCare + Why_StayFam + Why_StayHome,
#                    data=d,
#                controls=ctree_control(minbucket=.01*nrow(d)))
# 
# plot(mod.VPK7)
# 
# ####
# mod.VPK8=ctree(Why_AlternCare ~  MaritalStatus + HouseHoldIncome + NumChildrenEnteringKG + WillingToSpend2 + PrimaryTransportationMode + Trbl_NoChlgns,
#                    data=d,
#                controls=ctree_control(minbucket=.01*nrow(d)))
# 
# plot(mod.VPK8)
# 
# ####
# mod.VPK9=ctree(Why_AlternCare ~  MaritalStatus + HouseHoldIncome + NumChildrenEnteringKG + WillingToSpend2 + PrimaryTransportationMode + Trbl_NoPubSchls  + Trbl_NoPvtSchls + Trbl_Expnsv + Trbl_Location + Trbl_Other + Trbl_NoChlgns + Trbl_Quality + Trbl_NoFit + Trbl_Comfort + Trbl_NoHours + Trbl_Language,
#                    data=d,
#                controls=ctree_control(minbucket=.01*nrow(d)))
# 
# plot(mod.VPK9)
# 
# ####
# mod.VPK10=ctree(d$Why_StayHome ~  MaritalStatus + HouseHoldIncome + NumChildrenEnteringKG + WillingToSpend2 + PrimaryTransportationMode + Trbl_NoPubSchls  + Trbl_NoPvtSchls + Trbl_Expnsv + Trbl_Location + Trbl_Other + Trbl_NoChlgns + Trbl_Quality + Trbl_NoFit + Trbl_Comfort + Trbl_NoHours + Trbl_Language,
#                    data=d,
#                controls=ctree_control(minbucket=.01*nrow(d)))
# 
# plot(mod.VPK10)
# 
# 
# ####
# mod.VPK11=ctree(d$Why_StayFam ~  MaritalStatus + HouseHoldIncome + NumChildrenEnteringKG + WillingToSpend2 + PrimaryTransportationMode + Trbl_NoPubSchls  + Trbl_NoPvtSchls + Trbl_Expnsv + Trbl_Location + Trbl_Other + Trbl_NoChlgns + Trbl_Quality + Trbl_NoFit + Trbl_Comfort + Trbl_NoHours + Trbl_Language,
#                    data=d,
#                controls=ctree_control(minbucket=5))
# 
# plot(mod.VPK11)


#Keep Mod 7 & Mod 9 (below)

mod.VPK7=ctree(Trbl_NoChlgns ~  MaritalStatus + HouseHoldIncome + NumChildrenEnteringKG + WillingToSpend2 + PrimaryTransportationMode + Why_AlternCare + Why_StayFam + Why_StayHome,
                   data=d,
               controls=ctree_control(minbucket=.01*nrow(d)))

plot(mod.VPK7)

mod.VPK9=ctree(Why_AlternCare ~  MaritalStatus + HouseHoldIncome + NumChildrenEnteringKG + WillingToSpend2 + PrimaryTransportationMode + Trbl_NoPubSchls  + Trbl_NoPvtSchls + Trbl_Expnsv + Trbl_Location + Trbl_Other + Trbl_NoChlgns + Trbl_Quality + Trbl_NoFit + Trbl_Comfort + Trbl_NoHours + Trbl_Language,
                   data=d,
               controls=ctree_control(minbucket=.01*nrow(d)))
plot(mod.VPK9)


mod.VPK12=ctree(Why_AlternCare ~  MaritalStatus + HouseHoldIncome + NumChildrenEnteringKG + WillingToSpend2 + PrimaryTransportationMode + Trbl_NoPubSchls  + Trbl_NoPvtSchls + Trbl_Expnsv + Trbl_Location + Trbl_Other + Trbl_NoChlgns + Trbl_Comfort + Trbl_NoHours + Trbl_Language,
                   data=d,
               controls=ctree_control(minbucket=.01*nrow(d)))
plot(mod.VPK12)

###########
d$Why_AlternCare2=as.factor(d$Why_AlternCare)
mod.VPK9_2=ctree(Why_AlternCare2 ~  MaritalStatus + HouseHoldIncome + NumChildrenEnteringKG + WillingToSpend2 + PrimaryTransportationMode + Trbl_NoPubSchls  + Trbl_NoPvtSchls + Trbl_Expnsv + Trbl_Location + Trbl_Other + Trbl_NoChlgns + Trbl_Quality  Trbl_NoFit + Trbl_Comfort + Trbl_NoHours + Trbl_Language,
                data=d,
                controls=ctree_control(minbucket=.01*nrow(d)))





# png("out9_2.png",1200,1200)
# plot(mod.VPK10)
# dev.off()

```
  
      

```{r}



temp <- as.POSIXct( unique(data$VPKCombined$ProviderOpensAM) , format = "%H:%M" , tz = "GMT")

t=table(data$VPKCombined$ProviderOpensAM, data$VPKCombined$ProviderClosesPM, useNA = "always")

OpenOrder=unique(data$VPKCombined$ProviderOpensAM)[order(as.POSIXct(unique(data$VPKCombined$ProviderOpensAM) , format = "%H:%M" , tz = "GMT") , decreasing = FALSE )]
ClosedOrder=unique(data$VPKCombined$ProviderClosesPM)[order(as.POSIXct( unique(data$VPKCombined$ProviderClosesPM) , format = "%H:%M" , tz = "GMT"), decreasing = FALSE )]


    png("ProviderOpenAM_Revised.png")
       par(mar = c(12.1, 4.1, 4.1, 2.1))
        par(las=2)
        t=table(data$VPKCombined$ProviderOpensAM)
        b=barplot(t[OpenOrder], ylim=c(0,max(t)*1.2))
        text(b, t[OpenOrder] + max(t)*0.1, t)
      dev.off()

  png("ProviderClosedPM_Revised.png")

   par(mar = c(12.1, 4.1, 4.1, 2.1))
        par(las=2)
        t=table(data$VPKCombined$ProviderClosesPM)
        b=barplot(t[ClosedOrder], ylim=c(0,max(t)*1.2))
        text(b, t[ClosedOrder] + max(t)*0.1, t)
      dev.off()
```


```{r}

round(prop.table(table(data$VPKCombined$Education[(data$VPKCombined$Trbl_NoFit==0 & data$VPKCombined$Trbl_Quality==0)])),2)
round(prop.table(table(data$VPKCombined$Education[(data$VPKCombined$Trbl_NoFit==1 & data$VPKCombined$Trbl_Quality==1)])),2)


t1=round(prop.table(table(d$HouseHoldIncome[(d$Trbl_NoFit==0 & d$Trbl_Quality==0)])),2)
t2=round(prop.table(table(d$HouseHoldIncome[(d$Trbl_NoFit==1 & d$Trbl_Quality==1)])),2)

View(cbind(
"NoFitIssue"=round(prop.table(table(d$Education[(d$Trbl_NoFit==0 )])),2),
"NotreportedFit"=round(prop.table(table(d$Education[(d$Trbl_NoFit==1 )])),2),
"Diff"=round(prop.table(table(d$Education[(d$Trbl_NoFit==0 )])),2)-
round(prop.table(table(d$Education[(d$Trbl_NoFit==1 )])),2)))



round(prop.table(table(d$HouseHoldIncome[(d$Trbl_NoFit==0 )])),2)
round(prop.table(table(d$HouseHoldIncome[(d$Trbl_NoFit==1 )])),2)


# Fit/Quality problems in finding child care by education
prop.table(table(d$Trbl_NoFit==1,d$Education), margin=2)
prop.table(table(d$Trbl_Quality==1,d$Education), margin=2)
prop.table(table((d$Trbl_Quality==1 & d$Trbl_NoFit==1) ,d$Education), margin=2)

# Fit/Quality problems in finding child care by income
prop.table(table(d$Trbl_NoFit==1,d$HouseHoldIncome), margin=2)
prop.table(table(d$Trbl_Quality==1,d$HouseHoldIncome), margin=2)
prop.table(table((d$Trbl_Quality==1 & d$Trbl_NoFit==1),d$HouseHoldIncome), margin=2)


```



```{r}

##mod 7
round(prop.table(table("NoChlgns"=d$Trbl_NoChlgns[d$Why_AlternCare2==0], d$HHISplit[d$Why_AlternCare2==0]),margin=2),3)*100


##mod 9
round(prop.table(table("AltCare"=d$Why_AlternCare2, d$Trbl_NoChlgns),margin=2),3)*100

round(prop.table(table("AltCare"=d$Why_AlternCare2[d$Trbl_NoChlgns==0], d$HHISplit[d$Trbl_NoChlgns==0]),margin=2),3)*100


round(prop.table(table("AltCare"=d$Why_AlternCare2[d$Trbl_NoChlgns==0 & d$HHISplit==TRUE], d$Trbl_NoPubSchls[d$HHISplit==TRUE & d$Trbl_NoChlgns==0]),margin=2),3)*100

round(prop.table(table("AltCare"=d$Why_AlternCare2[d$Trbl_NoChlgns==0 & d$HHISplit==FALSE], d$Trbl_NoHours[d$HHISplit==FALSE & d$Trbl_NoChlgns==0]),margin=2),3)*100


round(prop.table(table("AltCare"=d$Why_AlternCare2[d$Trbl_NoChlgns==0 & d$HHISplit==FALSE & d$Trbl_NoHours==0], d$Trbl_NoFit[d$Trbl_NoChlgns==0 & d$HHISplit==FALSE & d$Trbl_NoHours==0]),margin=2),3)*100


round(prop.table(table("AltCare"=d$Why_AlternCare2[d$Trbl_NoChlgns==0 & d$HHISplit==FALSE & d$Trbl_NoHours==0 & d$Trbl_NoFit==0], d$Trbl_Quality[d$Trbl_NoChlgns==0 & d$HHISplit==FALSE & d$Trbl_NoHours==0 & d$Trbl_NoFit==0]),margin=2),3)*100


```




